---
title: "DATA 607 - Week 2 SQL & R"
author: "Brian K. Liles"
date: "February 10, 2018"
output:
  html_document: default
  word_document: default
---

#Synopsis

In this week's assignment six movie releases from 2017 served as the root of the SQL database **cinema** created in MySQL Workbench based on the following films:

*   + Star Wars: The Last Jedi
    + Jumanji: Welcome to the Jungle
    + Beauty and the Beast
    + Wonder Woman
    + Spider-Man Homecoming
    + The Fate of the Furious

##Data Dictionary

The **cinema** database is comprised of 17 varibles which are listed below; data provided from Wikipedia, IMDB, Rotten Tomatoes.

*   + **MID** - Movie ID variable
    + **Title** - Movie Title
    + **Descr** - Brief description
    + **Dir** - Director
    + **Dist** - Movie Studio Disbriutor
    + **Budget** - Movie Budget
    + **BoxOffice** - Box Office Profits
    + **RD_Month** - Release Date (Month)
    + **RD_Day** - Release Date (Day)
    + **RD_Year** - Release Date (Year)
    + **Rating** - Movie Rating (i.e. PG, PG-13)
    + **LengthHr** - Movie Length (Hour)
    + **LengthMin** - Movie Length (Minutes)
    + **IMDB** - IMDB Movie Rating
    + **Metascore** - Metascore Rating
    + **Tomato** - Rotten Tomatoes Rating
    + **AudScore** - Audience Score 
    
##MySQL Workbench

Data was exported from the **movies** table from MySQL Workbench to a csv file. In addition to the export the **DESCRIBE (TABLE)** statement output was also exported to a csv file. 

Load the tidyverse package

```{r}
library(tidyverse)
#import the describe csv file
describe <- read_csv("C:\\Users\\lizza\\Documents\\CUNY - Data Analytics\\DATA 607_Revised\\Week 2\\Assignment\\Cinema_Describe.csv")
```

By displaying the **describe** tibble we see that MID is considered a **primary key** which means it can't be duplicated; primary key's play an important role in databases as they allow several tables to link by SQL joins.

```{r}
#display the describe table
describe
```

##Import the Data

```{r}
#import the cinema csv
cinema <- read_csv("C:\\Users\\lizza\\Documents\\CUNY - Data Analytics\\DATA 607_Revised\\Week 2\\Assignment\\Cinema.csv")
glimpse(cinema)
```

View the names of the **cinema** tibble
```{r}
names(cinema)
```

Using the **tidyverse** rename the following imported variables: MID, Title, Descr, Dir, Dist, AudScore.

```{r}
#rename variables
cinema_new <- rename(cinema,MovieID=MID,MovieTitle=Title,Description=Descr,Director=Dir,Distributor=Dist,AudienceScore=AudScore)
#display the column names
colnames(cinema_new)
```

View the **cinema_new** tibble

```{r}
print(cinema_new)
```

Discover the film with the highest budget

```{r}
cinema_new %>% arrange(desc(Budget)) %>% select(MovieTitle,Budget)
```

Based on the findings we can see that **Jumanji:Welcome to the Jungle** had the lowest budget with 100 million versus **The Fate of the Furious** with the largest budget of 250 million dollars.

Discover the film that grossed the most at the box office

```{r}
cinema_new %>% arrange(desc(BoxOffice)) %>% select (MovieTitle,BoxOffice)
```

Based on the findings we can see that **Star Wars:The Last Jedi** was the winner of the 6 films by earning a little over 1.3 billion dollars while **Wonder Woman** earned over 820 million at the lowest slot.

Even though it is clear that **Star Wars: The Last Jedi** grossed the most at the box office we should look at the profits based on the budget;create a variable entitled **Profits**

```{r}
cinema_new %>% mutate(Profits = BoxOffice - Budget) %>% arrange (desc(Profits)) %>% select (MovieTitle,Profits)
```

Based on our findings **Star Wars: The Last Jedi** garnered the highest profits at the box office.

Next, create a new variable entitled **Ratio** which divides the **Profits** variable over the **BoxOffice** variable.

```{r}
cinema_new %>% mutate(Profits = BoxOffice - Budget) %>% mutate(Ratio = Profits / BoxOffice) %>% arrange (desc(Ratio)) %>% select(MovieTitle,Ratio)
```

Based on our findings **Jumanji: Welcome to the Jungle** was the most successful film out of the six.

Next, based on the **LengthHr** and **LengthMin** we will discover the longest movie out of the six by creating a variable entitled **TotalTime**.

```{r}
#multiply LengthHr by 60 to convert hours to minutes
cinema_new %>% mutate(TotalTime = (LengthHr * 60) + LengthMin) %>% arrange(desc(TotalTime)) %>% select (MovieTitle,TotalTime)
```

Based on our findings **Star Wars: The Last Jedi** was the longest film at 152 minutes while **Jumanji: Welcome to the Jungle** was the shortest with 119 minutes.

##Ratings
###IMDB

IMDB has become one of the leading sites that provide the public with ratings for television and film. Next, we will summarize the IMDB variable, plot the films and see which one had the highest ranking.

```{r}
summary(cinema_new$IMDB)
```

```{r}
cinema_new %>% arrange (desc(IMDB)) %>% select (MovieTitle,IMDB)
```

```{r}
barplot (cinema_new$IMDB, main = "IMDB Rankings of 2017 Films",xlab="Movie Titles", ylab="IMDB Rating", col= c("lightcyan","mistyrose","beige","yellow","red","green"), names.arg=c("Star","Jumanji","Beauty","Wonder","Spider-Man","Furious"))
```

Based on the findings, **Star Wars, Wonder Woman** and **Spider-Man** all received an IMDB score of 7.5 while **Fate of the Furious** had a score of 6.7.

###Metascore

Similar to **IMDB**, **Metascore** provides a ranking score for television, film and also video games. 

```{r}
summary(cinema_new$Metascore)
```

```{r}
cinema_new %>% arrange(desc(Metascore)) %>% select(MovieTitle,Metascore)
```

```{r}
barplot (cinema_new$Metascore, main = "Metascore Rankings of 2017 Films",xlab="Movie Titles", ylab="Metascore Rating", col= c("lightcyan","mistyrose","beige","yellow","red","green"), names.arg=c("Star","Jumanji","Beauty","Wonder","Spider-Man","Furious"))
```

**Star Wars** remains at the top, but this time without **Wonder Woman** and **Spider-Man**. **Jumanji** and **Fate of the Furious** are at the bottom of the **Metascore** rankings.

###Rotten Tomatoes & Audience Score

One of the newest ranking sites is **Rotten Tomatoes** which has their own system as well as an **AudienceScore** which we will also explore.

```{r}
summary(cinema_new$Tomato)
```

```{r}
cinema_new %>% arrange(desc(Tomato)) %>% select(MovieTitle,Tomato)
```

```{r}
barplot (cinema_new$Tomato, main = "Rotten Tomatoes Rankings of 2017 Films",xlab="Movie Titles", ylab="Rotten Tomatoes Rating", col= c("lightcyan","mistyrose","beige","yellow","red","green"), names.arg=c("Star","Jumanji","Beauty","Wonder","Spider-Man","Furious"))
```

With a narrow lead **Wonder Woman** and **Spider-Man** got the edge over **Star Wars** while it seems the **Fast & Furious** movies need to pump the brakes on new films.

```{r}
summary(cinema_new$AudienceScore)
```

```{r}
cinema_new %>% arrange (desc(AudienceScore)) %>% select(MovieTitle,AudienceScore)
```

```{r}
barplot (cinema_new$AudienceScore, main = "Audience Score (Rotten Tomatoes) of 2017 Films",xlab="Movie Titles", ylab="Audience Score Rating", col= c("lightcyan","mistyrose","beige","yellow","red","green"), names.arg=c("Star","Jumanji","Beauty","Wonder","Spider-Man","Furious"))
```

Here we see the audience ratings, and they were not pleased with **Star Wars** but welcomed **Jumanji, Wonder Woman, and Spider-Man**

#Conclusion

Based on the data we had to work with its clear there are endless possibilities if granted access to a full database where one could explore movie genres, movie releases by year, or delve deeper in directors and distribuors.
